Enhancing instance-level constrained clustering through differential evolution

### 2019-06-26 Papers/SHADE/paperSHADE.tex
Empezada memoria TFM
Nuevos capitulos en memoria y correcciones en papers

\begin{abstract}
Clustering has always been a powerful tool in knowledge discovery. Traditionally unsupervised, it received renewed attention when it was shown to produce better results when provided with new types of information, thus leading to a new kind of semi-supervised learning: constrained clustering. This technique is a generalization of traditional clustering that considers additional information encoded by constraints. Constraints can be given in the form of instance-level must-link and cannot-link constraints, in which this paper focuses on. We propose the first application of Differential Evolution to the constrained clustering problem, which has proven to produce a better exploration-exploitation trade-off when comparing with previous approaches. We will compare the results obtained by this proposal to those obtained by previous nature-inspired techniques and by some of the state-of-the-art algorithms on 25 datasets with incremental levels of constraint-based information, supporting our conclusions with the aid of Bayesian statistical tests.
\end{abstract}

@@

We can define partitional clustering as the task of grouping the instances of a dataset into $k$ clusters, so that new information can be extracted from them. A dataset $X$ is composed of $n$ instances, each one of them described by $d$ features. More formally, $X = \{x_1, \cdots, x_n\}$, with the $i$th instance noted as $x_i = (x_{[i,1]}, \cdots, x_{[i,u]})$. A typical clustering algorithm assigns a class label $l_i$ to each instance $x_i \in X$. As a result, we obtain the set of labels $L = \{l_1, \cdots, l_n\}$, with $l_i \in \{1, \cdots, k\}$, that effectively splits $X$ into $k$ non-overlapping clusters $c_i$ to form a partition called $C$. The criterion used to assign an instance to a given cluster is the similarity to the rest of elements in that cluster, and the dissimilarity to the rest of instances of the dataset, which can be obtained with some kind of distance measurement $d(\cdot, \cdot)$ \cite{jain1999data}.

@@

		$S_{CR} \leftarrow S_F \leftarrow \emptyset$\\
		$P_G \leftarrow$ \Sort{$P_G$} \tcp{Sort the population}
		\tcp{Apply an LS procedure to the elite of the population}
		\LocalSearch{$\{p_{[1,G]}, \cdots, p_{[p_{e},G]}\}$}\\
		\For{$i \in [1,p_{size}]$}{
			$r_i \leftarrow$ randInt$[1,|M|]$\\
			$CR_{[i,G]} \leftarrow$ randn$_i(M_{[CR,r_i]}, 0.1)$\\
			$F_{[i,G]} \leftarrow$ randc$_i(M_{[F,r_i]}, 0.1)$\\
			$p_{[i,G]} \leftarrow$ rand$[N/2, 0.2]$\\
			\tcp{Generate trial vector $t_{[i,G]}$ by current-to-$p$best/1/bin}

@@

		$s_i \leftarrow $ Select random object from $S$\\
		\tcp{Random shuffle labels set}
		$RSL \leftarrow $ \RandomShuffle{$\{1,\cdots,K\}$}\\
		\For{$l \in RSL$}{
			$S^\prime \leftarrow S$\\
			\tcp{Move object $s_i$ to the cluster associated with label $l$}
			$S^\prime$\texttt{[}$s_i$\texttt{]} $\leftarrow l$\\
			\If{$f(S^\prime) < f(S)$}{
				$S \leftarrow S^\prime$\\
				$improvement \leftarrow$ \texttt{true} \\
			}
		}
	}
	\BlankLine
	\KwRet ($S$)

@@

\begin{table}[!h]
	\centering
	\setlength{\tabcolsep}{7pt}
	\renewcommand{\arraystretch}{1.2}
	%\begin{adjustwidth}{-1in}{-1in}
	\resizebox{\textwidth}{!}{
		\begin{tabular}{l ccc c ccc}
			\hline
			\multirow{2}{*}{Dataset} &
			\multicolumn{3}{c}{BRKGA$_{CC}$} &&  \multicolumn{3}{c}{SHADE$_{CC}$} \\
			\cline{2-4} \cline{6-8}
			& ARI & Unsat(\%) & Time(s) && ARI & Unsat(\%) & Time(s) \\
			\hline
			Breast Cancer & 0.807 & 9.086 & 4061.162 && \textbf{0.977} & 0.891 & 3984.370 \\
			Bupa & 0.936 & 2.975 & 1529.151 && \textbf{0.993} & 0.281 & 1508.772 \\
			Circles & 0.935 & 2.814 & 1180.268 && \textbf{1.000} & 0.000 & 1170.452 \\
			Ecoli & 0.028 & 25.716 & 1267.419 && \textbf{0.047} & 21.466 & 1277.903 \\
			Glass & 0.038 & 23.610 & 766.861 && \textbf{0.082} & 16.412 & 796.162 \\
			Haberman & 0.943 & 2.073 & 1266.131 && \textbf{0.997} & 0.138 & 1235.441 \\
			Hayesroth & \textbf{0.548} & 6.331 & 554.339 && 0.531 & 7.056 & 540.585 \\
			Ionosphere & 0.902 & 4.314 & 2051.251 && \textbf{0.993} & 0.258 & 2019.558 \\
			Iris & \textbf{0.691} & 5.241 & 513.675 && 0.553 & 6.528 & 488.262 \\
			Led7Digit & \textbf{0.007} & 13.653 & 2053.264 && 0.005 & 12.170 & 2066.223 \\
			Monk2 & 0.855 & 6.025 & 2055.425 && \textbf{0.991} & 0.379 & 2013.468 \\
			Moons & 0.937 & 2.689 & 1177.848 && \textbf{0.995} & 0.158 & 1157.222 \\
			Movement Libras & \textbf{0.006} & 8.381 & 1654.159 && 0.004 & 7.676 & 1645.174 \\
			Pima & 0.753 & 11.544 & 4855.446 && \textbf{0.899} & 4.419 & 4820.684 \\
			Saheart & 0.881 & 5.250 & 2402.513 && \textbf{0.981} & 0.706 & 2353.931 \\
			Sonar & 0.992 & 0.186 & 1138.769 && \textbf{1.000} & 0.000 & 1127.020 \\
			Soybean & 0.416 & 0.000 & 171.281 && \textbf{0.563} & 0.000 & 153.406 \\
			Spectfheart & 0.953 & 1.677 & 1602.229 && \textbf{1.000} & 0.000 & 1526.925 \\
			Spiral & 0.921 & 3.604 & 1178.580 && \textbf{0.998} & 0.091 & 1155.194 \\
			Tae & 0.351 & 9.807 & 527.292 && \textbf{0.392} & 9.419 & 508.610 \\
			Vehicle & \textbf{0.144} & 28.781 & 5189.066 && 0.044 & 29.008 & 5284.288 \\
			Vowel & 0.002 & 14.682 & 5450.331 && \textbf{0.003} & 13.766 & 5514.052 \\
			Wdbc & 0.823 & 8.281 & 4019.683 && \textbf{0.972} & 1.068 & 3983.564 \\
			Zoo & \textbf{0.176} & 5.429 & 335.715 && 0.153 & 4.191 & 320.186 \\
			\hline
			Mean & 0.602 & 8.086 & 1895.671 && \textbf{0.647} & 5.443 & 1880.921 \\
			\hline

>>>>

\begin{abstract}
Clustering has always been a powerful tool in knowledge discovery. Traditionally unsupervised, it received renewed attention when it was shown to produce better results when provided with new types of information, thus leading to a new kind of semi-supervised learning: constrained clustering. This technique is a generalization of traditional clustering that considers additional information encoded by constraints. Constraints can be given in the form of instance-level must-link and cannot-link constraints, which this paper focuses on. We propose the first application of Differential Evolution to the constrained clustering problem, which has proven to produce a better exploration-exploitation trade-off when comparing with previous approaches. We will compare the results obtained by this proposal to those obtained by previous nature-inspired techniques and by some of the state-of-the-art algorithms on 25 datasets with incremental levels of constraint-based information, supporting our conclusions with the aid of Bayesian statistical tests.
\end{abstract}

@@

We can define partitional clustering as the task of grouping the instances of a dataset into $k$ clusters, so that new information can be extracted from them. A dataset $X$ is composed of $n$ instances, each one of them described by $u$ features. More formally, $X = \{x_1, \cdots, x_n\}$, with the $i$th instance noted as $x_i = (x_{[i,1]}, \cdots, x_{[i,u]})$. A typical clustering algorithm assigns a class label $l_i$ to each instance $x_i \in X$. As a result, we obtain the set of labels $L = \{l_1, \cdots, l_n\}$, with $l_i \in \{1, \cdots, k\}$, that effectively splits $X$ into $k$ non-overlapping clusters $c_i$ to form a partition called $C$. The criterion used to assign an instance to a given cluster is the similarity to the rest of elements in that cluster, and the dissimilarity to the rest of instances of the dataset, which can be obtained with some kind of distance measurement $d(\cdot, \cdot)$ \cite{jain1999data}.

@@

		$S_{CR} \leftarrow S_F \leftarrow \emptyset$\\
		$P_G \leftarrow$ \Sort{$P_G$} \tcp{Sort the population}
		\tcp{Apply an LS procedure to the elite of the population}
		\LocalSearch{$\{p_{[1,G]}, \cdots, p_{[p_{e},G]}\}$}\\
		\For{$i \in [1,p_{size}]$}{
			$r_i \leftarrow$ randInt$[1,|M|]$\\
			$CR_{[i,G]} \leftarrow$ randn$_i(M_{[CR,r_i]}, 0.1)$\\
			$F_{[i,G]} \leftarrow$ randc$_i(M_{[F,r_i]}, 0.1)$\\
			$p_{[i,G]} \leftarrow$ rand$[N/2, 0.2]$\\
			Generate trial vector $t_{[i,G]}$ by current-to-$p$best/1/bin\\

@@

		$s_i \leftarrow $ Select random object from $S$\\
		\tcp{Random shuffle labels set}
		$RSL \leftarrow $ \RandomShuffle{$\{1,\cdots,K\}$}\\
		\For{$l \in RSL$}{
			$S^\prime \leftarrow S$\\
			\tcp{Move object $s_i$ to the cluster associated with label $l$}
			$S^\prime$\texttt{[}$s_i$\texttt{]} $\leftarrow l$\\
			\If{$f(S^\prime) < f(S)$}{
				$S \leftarrow S^\prime$\\
				$improvement \leftarrow$ \texttt{true} \\
			}
		}
	}
	\BlankLine
	\KwRet{$S$}

@@

\begin{table}[!h]
	\centering
	\setlength{\tabcolsep}{7pt}
	\renewcommand{\arraystretch}{1.2}
	%\begin{adjustwidth}{-1in}{-1in}
	\resizebox{\textwidth}{!}{
		\begin{tabular}{l ccc c ccc}
			\hline
			\multirow{2}{*}{Dataset} &
			\multicolumn{3}{c}{BRKGA$_{CC}$} &&  \multicolumn{3}{c}{SHADE$_{CC}$} \\
			\cline{2-4} \cline{6-8}
			& ARI & Unsat(\%) & Time(s) && ARI & Unsat(\%) & Time(s) \\
			\hline
			Appendicitis & 1.000 & 0.000 & 389.905 && 1.000 & 0.000 & 371.579 \\
			Breast Cancer & 0.807 & 9.086 & 4061.162 && \textbf{0.977} & 0.891 & 3984.370 \\
			Bupa & 0.936 & 2.975 & 1529.151 && \textbf{0.993} & 0.281 & 1508.772 \\
			Circles & 0.935 & 2.814 & 1180.268 && \textbf{1.000} & 0.000 & 1170.452 \\
			Ecoli & 0.028 & 25.716 & 1267.419 && \textbf{0.047} & 21.466 & 1277.903 \\
			Glass & 0.038 & 23.610 & 766.861 && \textbf{0.082} & 16.412 & 796.162 \\
			Haberman & 0.943 & 2.073 & 1266.131 && \textbf{0.997} & 0.138 & 1235.441 \\
			Hayesroth & \textbf{0.548} & 6.331 & 554.339 && 0.531 & 7.056 & 540.585 \\
			Ionosphere & 0.902 & 4.314 & 2051.251 && \textbf{0.993} & 0.258 & 2019.558 \\
			Iris & \textbf{0.691} & 5.241 & 513.675 && 0.553 & 6.528 & 488.262 \\
			Led7Digit & \textbf{0.007} & 13.653 & 2053.264 && 0.005 & 12.170 & 2066.223 \\
			Monk2 & 0.855 & 6.025 & 2055.425 && \textbf{0.991} & 0.379 & 2013.468 \\
			Moons & 0.937 & 2.689 & 1177.848 && \textbf{0.995} & 0.158 & 1157.222 \\
			Movement Libras & \textbf{0.006} & 8.381 & 1654.159 && 0.004 & 7.676 & 1645.174 \\
			Pima & 0.753 & 11.544 & 4855.446 && \textbf{0.899} & 4.419 & 4820.684 \\
			Saheart & 0.881 & 5.250 & 2402.513 && \textbf{0.981} & 0.706 & 2353.931 \\
			Sonar & 0.992 & 0.186 & 1138.769 && \textbf{1.000} & 0.000 & 1127.020 \\
			Soybean & 0.416 & 0.000 & 171.281 && \textbf{0.563} & 0.000 & 153.406 \\
			Spectfheart & 0.953 & 1.677 & 1602.229 && \textbf{1.000} & 0.000 & 1526.925 \\
			Spiral & 0.921 & 3.604 & 1178.580 && \textbf{0.998} & 0.091 & 1155.194 \\
			Tae & 0.351 & 9.807 & 527.292 && \textbf{0.392} & 9.419 & 508.610 \\
			Vehicle & \textbf{0.144} & 28.781 & 5189.066 && 0.044 & 29.008 & 5284.288 \\
			Vowel & 0.002 & 14.682 & 5450.331 && \textbf{0.003} & 13.766 & 5514.052 \\
			Wdbc & 0.823 & 8.281 & 4019.683 && \textbf{0.972} & 1.068 & 3983.564 \\
			Zoo & \textbf{0.176} & 5.429 & 335.715 && 0.153 & 4.191 & 320.186 \\
			\hline
			Mean & 0.602 & 8.086 & 1895.671 && \textbf{0.647} & 5.443 & 1880.921 \\
			\hline
